---
title: "2024-09-12"
format: docx
---


# Setup

```{r}
data(mtcars)
library(easystats)
library(tidyverse)
```


# First model

<https://stats-nutshell.netlify.app/050-regression1#first-model-one-metric-predictor>

```{r}
lm1_freq <- lm(mpg ~ hp, data = mtcars)
```


Extract parameters of the model:

```{r}
parameters(lm1_freq) |> 
  print_md()
```


```{r}
plot(parameters(lm1_freq),
     show_intercept = TRUE)

plot(parameters(lm1_freq))  # better without

```


# Exercise


Consider the dataset `penguins`. Compute a linear model with body mass as output variable (DV) and flipper length as input (IV). 

1. Report the coefficients and interpret them.
2. Plot the model and the coefficients.
3. Report the model fit (R squared).
4. BONUS: `predict()` the weight of an average flipper-seized animal.  Check out the internet for examples of how to do so in case you need support.




# Solution

## Setup

```{r}
library(tidyverse)
library(easystats)
library(ggpubr)

# import data:
penguins <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")
```



## Let's go


```{r}
lm1 <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```



Plot the model:


```{r}
plot(estimate_relation(lm1))
```



Alternative plotting method:


```{r}
ggscatter(penguins,
          x = "flipper_length_mm",
          y = "body_mass_g",
          add ="reg.line")
```



Coefficients (parameters):


```{r}
parameters(lm1)
```



Plot the coefficients:

```{r}
plot(parameters(lm1))
```



Model fit (explained variance by model):



```{r}
r2(lm1)
```



Predict weight of average animal:

```{r}
penguins |> 
  summarise(flipper_length_mm_avg = mean(flipper_length_mm, na.rm = TRUE))
```



```{r}
predict(lm1, newdata = data.frame(flipper_length_mm = 200))
```

Around 4 kgs.


Center the data:


```{r}
penguins_c <-
  penguins |> 
  mutate(flipper_length_mm_c = center(flipper_length_mm))
```


Now the mean value is (nearly) zero:

```{r}
mean(penguins_c$flipper_length_mm_c, na.rm = TRUE)
```


Run the model again:

```{r}
lm2 <- lm(body_mass_g ~ flipper_length_mm_c, data = penguins_c)

parameters(lm2)
```


