---
title: "My Quarto document!!!"
format:
  html
  # pdf: 
  #   keep-tex: true
---


# Table One example


Load the needed R packages:


R code must be explicitly marked as code, use "R code chunks",
shortcut Command-Option-I, or hit the "+C" green button


```{r}
library(tableone)
library(DataExplorer)
```


Now load the dataset:

```{r}
penguins <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")  # hit command-enter
```



# DataExplorer


Hit Control-Enter to run some code line:

```{r}
plot_histogram(penguins)
```




# TableOne

```{r}
CreateTableOne(data = penguins) |>
  kableone()  # adorns styling for html/pdf/..
```


# THE PIPE

"And then" command: vertical bar + greater-than "|" + ">" -
is called the PIPE
and can be read as "and then do ..."
the shortcut is **"Control+Shift+M"**.



# Exercise

Show a grouped (and beautiful) table for all metric variables in the `penguins` dataset.
Note that you'll need to R packages: tidyverse and easystats.



```{r}
library(easystats)
library(tidyverse)
penguins |> 
  group_by(sex) |> 
  describe_distribution() |> 
  print_md()
```






# Ex: Visualize the mtcars dataset

```{r}
data(mtcars)
```


Use the tools presented here to visualize variables of interest of the mtcars dataset. In addition, provided plots that investigate bivariate associations (i.e., potential correlations between two variables).

See here for details:

<https://stats-nutshell.netlify.app/030-eda#exercises>

