---
title: "My title"
author: "Me"
date: today
format: html
---

# That's how to do it

1. Start the libraries
2. Import the data
3. Create the table with the distributions.

Then hit on "render".

# Start libraries

```{r}
#| message: false  # suppresses messages
#| warning: false  # suppresses warnings
library(tidyverse)
library(easystats)
library(ggpubr)
```


# Import data

```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")
```


# Describe distribution


```{r}
d |> 
  describe_distribution() |> 
  print_md()  # make a more styled table
```


# Describe non-numeric data

How many female and male aninals are in the data set?


```{r}
d |> 
  data_tabulate(select = sex)
```



```{r}
d |> 
  select(sex) |> 
  data_tabulate() |> 
  print_md()
```

Empty text/string/character values:

```{r}
my_dummy_var <- NA
```


# Get rid of empty strings


```{r}
d_no_na2 <- 
d |> 
  drop_na() |> 
  # now empty strings:
  filter(sex != "") # ! =
```



Exercise: Filter the data set such that only female animals remain.


```{r}
d <- d_no_na2
d %>%
  drop_na() %>%
  dplyr::filter(sex != "")
```



# Data vis


## Numeric X, Numeric Y

```{r}
ggscatter(d_no_na2,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "sex")
```



More on `ggscatter`:


```{r}
ggscatter(d_no_na2, 
           x = "bill_length_mm",
           y = "body_mass_g",     
           color = "species", 
          title = "penguin2.0",
          facet.by = "sex",
        add = "reg.line") +
  scale_color_okabeito()
```



## Numeric X

Visualize the distribution of the body weight of the penguins.

```{r}
gghistogram(data = d,
            x = "body_mass_g")
```

More histograms:

```{r}
gghistogram(d_no_na,
            x = "body_mass_g",
            facet.by = "species",
            xlab = "Body mass [g]",
            color = "black",
            fill = "blue",
            bins = 50)
```


```{r}
gghistogram(d_no_na,
            x = "body_mass_g",
            fill = "species",
            shape = "species",
            conf.int = TRUE,
            ellipse = TRUE,
            cor.coef = TRUE)
```
```{r}
ggdensity(d_no_na,
            x = "body_mass_g",
         # facet.by = "species",
          fill = "species")
```

## Metric Y, nominal X variable

Compare the weight between species

### Boxplot



```{r}
ggboxplot(penguins,
          x = "species",
          y = "body_mass_g",
          notch = TRUE,
          add = "mean",
          add.params = 
            list(color = okabeito_colors()[1],
                 size = 1)
          )
```

### Violinplot

Try a "violinplot" and see if you can make sense out of it.




```{r}
#| warning: false
ggviolin(penguins,
          x = "species",
          y = "body_mass_g",
          notch = TRUE,
          add = "mean_se",
          add.params = 
            list(color = okabeito_colors()[1],
                 size = 1)
          )
```


# Inference!


Research question:

>   Male penguins are heavier than female ones *on average*.

More mathy: $\mu_M > \mu_F$

What's the DV and IV here?

- DV: weight
- IV: sex



Tilde: "~" -> Windows: Alt GR + "+" key, Mac: Option+n

```{r}
lm1 <- lm(body_mass_g ~ sex, data = d_no_na2)
parameters(lm1)
```

```{r}
library(rstanarm)
```

Bayes:

```{r}
lm1 <- stan_glm(body_mass_g ~ sex, data = d_no_na2)
parameters(lm1)
```

