---
title: "My title"
author: "Me"
date: today
format: html
---

# That's how to do it

1. Start the libraries
2. Import the data
3. Create the table with the distributions.

Then hit on "render".

# Start libraries

```{r}
#| message: false  # suppresses messages
#| warning: false  # suppresses warnings
library(tidyverse)
library(easystats)
library(ggpubr)
```


# Import data

```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")
```


# Describe distribution


```{r}
d |> 
  describe_distribution() |> 
  print_md()  # make a more styled table
```


# Describe non-numeric data

How many female and male aninals are in the data set?


```{r}
d |> 
  data_tabulate(select = sex)
```



```{r}
d |> 
  select(sex) |> 
  data_tabulate() |> 
  print_md()
```

Empty text/string/character values:

```{r}
my_dummy_var <- NA
```


# Get rid of empty strings


```{r}
d_no_na2 <- 
d |> 
  drop_na() |> 
  # now empty strings:
  filter(sex != "") # ! =
```



Exercise: Filter the data set such that only female animals remain.


```{r}
d %>%
  drop_na() %>%
  dplyr::filter(sex != "")
```



# Data vis


## Numeric X, Numeric Y

```{r}
ggscatter(d_no_na2,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "sex")
```



More on `ggscatter`:


```{r}
ggscatter(d_no_na2, 
           x = "bill_length_mm",
           y = "body_mass_g",     
           color = "species", 
          title = "penguin2.0",
          facet.by = "sex",
        add = "reg.line") +
  scale_color_okabeito()
```



## Numeric X

Visualize the distribution of the body weight of the penguins.

```{r}
gghi
```





