---
title: "20260922 - Introduction to statistical modeling"
format: docx
---


# Welcome!


# Quick start


Load R packages:

Use `Alt+Command+I` to *I*nsert a Code Chunk:

```{r}

```




```{r}
library(tidyverse)  
library(easystats)
```

Load data:


```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

head(d)  # see the first few rows, the "head" of the table
```


First Modelling:


```{r}
m1 <- lm(body_mass_g ~ bill_length_mm, data = d)  # Control+Enter

model_parameters(m1)
```

$R^2$: The (percentage) amount of explained variance of the model - "goodness" of the model.

```{r}
r2(m1)
```


Standardized coefficients:

```{r}
effectsize::standardize_parameters(m1, method = "refit")   # oder "basic"
```

Visualize the model:

```{r}
library(ggpubr)  #control+enter

ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "species",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          alpha = 0.6,
          xlab = "Schnabellänge (mm)",
          ylab = "Körpermasse (g)")
```




# Exercise

Choose and run your own model, just as above. Try to understand the result.


# Descriptives statistics

Control+Shift+I -- insert code chunk

Mind the pipe! "|>"

```{r}
d |>  # Control+Shift+M - the "pipe" - "and then do ..."
  select(bill_length_mm, body_mass_g, sex) |> 
  describe_distribution(by = "sex")  |> 
  print_md()
```


```{r}
ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "sex",
          add = "reg.line", # regression line
          shape = "sex")
```


# Exercise

Find a function to visualize the difference between multiple groups in one plot. Use the package `ggbpur`. Try it out!


```{r}
ggviolin(d, 
x = "sex", 
y = "body_mass_g", 
fill = "sex",
palette = c("#00AFBB", "#E7B800", "#FC4E07"),
add = c("boxplot", "mean"), 
add.params = list(fill = "white"))
```


