---
title: "20260922, 24 - Introduction to statistical modeling"
format: html
editor_options: 
  chunk_output_type: console
---


# Welcome!


# Quick start


Load R packages:

Use `Alt+Command+I` to *I*nsert a Code Chunk:

```{r}

```




```{r}
library(tidyverse)  
library(easystats)
```

Load data:


```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

head(d)  # see the first few rows, the "head" of the table
```


First Modelling:


```{r}
m1 <- lm(body_mass_g ~ bill_length_mm, data = d)  # Control+Enter

model_parameters(m1)
```

$R^2$: The (percentage) amount of explained variance of the model - "goodness" of the model.

```{r}
r2(m1)
```


Standardized coefficients:

```{r}
effectsize::standardize_parameters(m1, method = "refit")   # oder "basic"
```

Visualize the model:

```{r}
library(ggpubr)  #control+enter

ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "species",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          alpha = 0.6,
          xlab = "Schnabellänge (mm)",
          ylab = "Körpermasse (g)")
```




# Exercise

Choose and run your own model, just as above. Try to understand the result.


# Descriptives statistics

Control+Shift+I -- insert code chunk

Mind the pipe! "|>"

```{r}
d |>  # Control+Shift+M - the "pipe" - "and then do ..."
  select(bill_length_mm, body_mass_g, sex) |> 
  describe_distribution(by = "sex")  |> 
  print_md()
```


```{r}
ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "sex",
          add = "reg.line", # regression line
          shape = "sex")
```


# Exercise

Find a function to visualize the difference between multiple groups in one plot. Use the package `ggbpur`. Try it out!


```{r}
ggviolin(d, 
x = "sex", 
y = "body_mass_g", 
fill = "sex",
palette = c("#00AFBB", "#E7B800", "#FC4E07"),
add = c("boxplot", "mean"), 
add.params = list(fill = "white"))
```



# Modelling y ~ x + b

Is the weight of a penguin related to its bill length and to the sex of the animal?

```{mermaid}
flowchart LR
x --> y
b --> y
```


```{mermaid}
flowchart LR
bill_length --> body_weight
sex --> body_weight
```

# First: tidy up the data


```{r}
d |> 
   filter(sex == "female" | sex == "male" ) -> d2
```


same thing:

```{r}
d2 <- 
d |> 
   filter(sex == "female" | sex == "male")
```

# Model it

```{r}
m2 <- lm(body_mass_g ~ bill_length_mm + sex, data = d2)  # Control+Enter for execution

model_parameters(m2)
```

# Visualize it

Tip: To run all of the code in this file - hit `Control+Alt+R`.


```{r}
estimate_relation(m2) |> 
  plot(show_data = TRUE) +
  scale_color_okabeito() +
  scale_fill_okabeito() +
  theme_modern()
```

# Excercise


`y ~ x + b`

Find an example research question from our penguin data set,
where 

- y - is metric (outcome, DV)
- x - is metric (input)
- b - is binary (two groups) (input)

# Exercise

Research question:

>   Is the flipper length (DV) of a penguin related to its bill depth and the sex of the animal?


Compute the model and report its goodness ($R²$) - also the model parameters.


```{r}
m3 <- lm(flipper_length_mm ~ bill_depth_mm + sex, data = d2) 

model_parameters(m3)

summary(m3)

m3_r2 <- r2(m3)
```


$R^2$ was found to be `r r2(m3)$R2`.


The number is 73.

# Exercise

`y ~ b` 

Find a research question for this model (from the penguins data set).

>  Is bill length in penguins a function of sex?

>  Have male penguins a longer bill than female penguins?


this is the so-called *t-test* scenario.

```{r}
lm(bill_length_mm ~ sex, data = d2)
```





# Exercise

## Consider

>   Is the weight a function of the the species of the penguins?

Find the statistical taxonomy model and compute the parameters. And visualize it. And tell the model goodness (explained variance).

Bonus: How could we formulate the research question in a more precise way?


## Solution


prototype statistical taxonomy model:

`y ~ g`

with

- y : metric outcome
- g : nominal input variable (with 2 or more groups)

```{r}
m_weight_species <- lm(body_mass_g ~ species, data = d2)

model_parameters(m_weight_species)

r2(m_weight_species)

estimate_relation(m_weight_species) %>%
  plot(show_data = T) +
  theme_minimal()
```

Research question:

>   Is body mass a function of the species?

Research question more precise:

>   Are penguins of the Adelie species heavier than Gentoo or Chinstrap penguins?

>  ... on average ...

$\mu_{\text{Adelie}} \ne \mu_2$






# Bayes

Frequentist:

```{r}
lm(bill_length_mm ~ sex, data = d2)
```

Bayes:


```{r}
library(rstanarm)

m4 <- stan_glm(bill_length_mm ~ sex, data = d2)

model_parameters(m4)
```


# More models


## Interaction effect

`y ~ x + g + x:g`

synonymous:

`y ~ x * g`

compare to without interaction effect:


`y ~ x + g`


What's an interaction effect in intuitive terms?


sugar + stirring --> sweetness


rain + sun --> plant growth


Does the effect of X on y depends on g?

if the slopes are NOT parallel.


Research question for interaction effect (penguins data):


>   Does the effect of bill length on weight depends on sex of the penguin?


```{r}
m6 <- lm(body_mass_g ~ bill_length_mm * sex, data = d2)

model_parameters(m6)
```


Diagram for m6:


```{r}
estimate_relation(m6) |> 
  plot(show_data = TRUE) +
  scale_color_okabeito() +
  scale_fill_okabeito() +
  theme_modern()
```


# Ex


## Question

>   weight as a function of bill length and species

Please report:

1. statistical taxonomy model
2. compute it and report parameters 
3. report r2
4. show visualization

## Answer

1. statistical taxonomy model

`y ~ x* g`

`y ~ x + g + x:g`

2. compute it and report parameters 

```{r}
m_interact2 <- lm(body_mass_g ~ bill_length_mm * species, data = d2)
model_parameters(m_interact2)
```


3. report r2

```{r}
r2(m_interact2)
```

4. show visualization

```{r}
estimate_relation(m_interact2) |> 
  plot(show_data = TRUE) +
  theme_modern()
```

## Bonus - stability of coefficients

```{r}
m_bill_length <-  lm(body_mass_g ~ bill_length_mm, data = d2)
model_parameters(m_bill_length)
```

In this case, the coefficient `bill_length` is stable - but in other case, the coefficient might vary wildly!

## Bonus - visualize parameters


```{r}
m_interact2 |> 
  model_parameters() |> 
  plot()
```

# Centering the data

```{r}
d3 <- 
d2 |> 
  center(select = "bill_length_mm",
         append = TRUE)
```

```{r}
d2 |> 
  describe_distribution(select = "bill_length_mm")
```

avarage: approx. 44mm


# Ex

## Question

Please refit the model `m_interact2` with centered X (bill_length_mm) and compare the parameter values. Any difference?


## Solution

```{r}
m_interact3 <- lm(body_mass_g ~ bill_length_mm_c * species, data = d3)
model_parameters(m_interact3)

model_parameters(m_interact2)
```

# Standardize the data (z-transformation)

# Bayes compared to Frequentist

## Ex

Take `m_interact3` and run a Bayesian analysis with it.

tip:

- package `rstanarm`
- command `stan_glm` instead of `lm`
- the rest remains the same



