---
title: "20260922, 24 - Introduction to statistical modeling"
format: html
---


# Welcome!


# Quick start


Load R packages:

Use `Alt+Command+I` to *I*nsert a Code Chunk:

```{r}

```




```{r}
library(tidyverse)  
library(easystats)
```

Load data:


```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

head(d)  # see the first few rows, the "head" of the table
```


First Modelling:


```{r}
m1 <- lm(body_mass_g ~ bill_length_mm, data = d)  # Control+Enter

model_parameters(m1)
```

$R^2$: The (percentage) amount of explained variance of the model - "goodness" of the model.

```{r}
r2(m1)
```


Standardized coefficients:

```{r}
effectsize::standardize_parameters(m1, method = "refit")   # oder "basic"
```

Visualize the model:

```{r}
library(ggpubr)  #control+enter

ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "species",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          alpha = 0.6,
          xlab = "Schnabellänge (mm)",
          ylab = "Körpermasse (g)")
```




# Exercise

Choose and run your own model, just as above. Try to understand the result.


# Descriptives statistics

Control+Shift+I -- insert code chunk

Mind the pipe! "|>"

```{r}
d |>  # Control+Shift+M - the "pipe" - "and then do ..."
  select(bill_length_mm, body_mass_g, sex) |> 
  describe_distribution(by = "sex")  |> 
  print_md()
```


```{r}
ggscatter(d,
          x = "bill_length_mm",
          y = "body_mass_g",
          color = "sex",
          add = "reg.line", # regression line
          shape = "sex")
```


# Exercise

Find a function to visualize the difference between multiple groups in one plot. Use the package `ggbpur`. Try it out!


```{r}
ggviolin(d, 
x = "sex", 
y = "body_mass_g", 
fill = "sex",
palette = c("#00AFBB", "#E7B800", "#FC4E07"),
add = c("boxplot", "mean"), 
add.params = list(fill = "white"))
```



# Modelling y ~ x + b

Is the weight of a penguin related to its bill length and to the sex of the animal?

```{mermaid}
flowchart LR
x --> y
b --> y
```


```{mermaid}
flowchart LR
bill_length --> body_weight
sex --> body_weight
```

# First: tidy up the data


```{r}
d |> 
   filter(sex == "female" | sex == "male" ) -> d2
```


same thing:

```{r}
d2 <- 
d |> 
   filter(sex == "female" | sex == "male")
```

# Model it

```{r}
m2 <- lm(body_mass_g ~ bill_length_mm + sex, data = d2)  # Control+Enter for execution

model_parameters(m2)
```

# Visualize it

Tip: To run all of the code in this file - hit `Control+Alt+R`.


```{r}
estimate_relation(m2) |> 
  plot(show_data = TRUE) +
  scale_color_okabeito() +
  scale_fill_okabeito() +
  theme_modern()
```

# Excercise


`y ~ x + b`

Find an example research question from our penguin data set,
where 

- y - is metric (outcome, DV)
- x - is metric (input)
- b - is binary (two groups) (input)

# Exercise

Research question:

>   Is the flipper length (DV) of a penguin related to its bill depth and the sex of the animal?


Compute the model and report its goodness ($R²$) - also the model parameters.


```{r}
m3 <- lm(flipper_length_mm ~ bill_depth_mm + sex, data = d2) 

model_parameters(m3)

summary(m3)

m3_r2 <- r2(m3)
```


$R^2$ was found to be `r r2(m3)$R2`.


The number is 73.

# Exercise

`y ~ b` 

Find a research question for this model (from the penguins data set).






