---
title: "2025-09-16"
format: html
---


# Import data

## Importing data from the internet

```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

# head(d)  # see the first few rows, the "head" of the table
```


## Import data from your own project folder (own computer)


```{r}
penguins <- read.csv("penguins.csv")
```



# Starting packages

```{r}
library(easystats)
```


# Descritive stasts: Getting a table

```{r}
describe_distribution(d) |> print_md()
```


# Visualize models of the data


Taxonomy:

1. distribution
2. association (r)
3. differences between groups



## nice package for that: ggpubr

Add a code chunk and load this package:

Command+Option+I adds a code chunk
```{r}
library(ggpubr)
```


## Distributions

```{r}
gghistogram(penguins, 
            x = "body_mass_g",
            add = "mean"
            )

```

# Assocation

```{r}
ggscatter(data = d, 
          x = "bill_length_mm",
          y = "body_mass_g",
          add = "reg.line")
```


## Differences

```{r}
ggboxplot(penguins, 
          x = "sex",
          y = "body_mass_g")
```



## Exercise


```{r}
ggscatter(data = d,
x = "bill_length_mm", 
y = "bill_depth_mm",
add  = "reg.line",
color = "species")

```


```{r}
ggboxplot(d, 
          x = "sex",
          y = "flipper_length_mm")
```


# Regression with R


## numeric x, numeric y


```{r}
ggscatter(data = d, 
          x = "bill_length_mm",
          y = "body_mass_g",
          add = "reg.line")
```

lm: "linear model"
"y is a function of x" - function is defined in math
"~": is a function of "tilde" (left to the enter key on Windows, Mac: Option+n)

```{r}
lm(body_mass_g ~ bill_length_mm, data = d)
```

## Standardizing (z-transformation) the data 

- Helpful for interpretation.

centering: subtracting the mean value (in X) from each data point

z-standardizing: "normalizing" 
 - *centering*
 - standardizing the *scale* (putting the variance/standard devation to 1)

```{r}
penguins_standardized <- 
d |>  # "and then"  
  standardize(select = "bill_length_mm") 
```


```{r}
lm_body_mass_bill_lenth_z <- 
lm(body_mass_g ~ bill_length_mm, data = penguins_standardized)

lm_body_mass_bill_lenth_z
```


## Exercise

Run a regression:

y: body mass g
x flipper length - z-transformed

And interpret the coefficients



```{r}
penguins_standardized_flipper <- 
d |>  # "and then"  
  standardize(select = "flipper_length_mm") 
```


```{r}
lm_body_mass_flipper_length_z <- 
lm(body_mass_g ~ flipper_length_mm, data = penguins_standardized_flipper)

lm_body_mass_flipper_length_z
```

## p-value and other inferential statistics

with beautification: 

```{r}
parameters(lm_body_mass_flipper_length_z) |> 
  print_md()
```

significant effect!


## R-Squared

The amount (%) of explained variance by your model.
The higher the better.


```{r}
r2(lm_body_mass_flipper_length_z)
```


## x: two groups, y: numeric



```{r}
ggboxplot(d, 
          x = "sex",
          y = "flipper_length_mm")
```




## Data cleansing


```{r}
d_filtered <- 
d |>  # and then
  filter(sex == "male" | sex == "female")
```


## Exercise: Regression

with filtered data

x: the grouping variable: sex
y: the weight

Plot:

```{r}
ggboxplot(d_filtered, 
        x = "sex",
        y = "body_mass_g")
```


```{r}
lm_weight_by_sex <-
  lm(body_mass_g ~ sex, data = d_filtered)

parameters(lm_weight_by_sex)
```

## t-Test is the same, more or less

```{r}
t.test(body_mass_g ~ sex, data = d_filtered)
```

