---
title: "20260223_25"
format: html
---


Hi there!
Let me jot down some notes...


```{r}
1+1
```

# Import data

```{r}
d <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/palmerpenguins/penguins.csv")

head(d)  # see the first few rows, the "head" of the table
```

Import data from your own machine:

Either use the button "import dataset" or use the `read.csv` command.


```{r}
#penguins <- read.csv("~/Downloads/penguins.csv")
```



# Simple model: `y ~ b`

"Y as a function of of a binary input variable (two groups)".

> Is the weight of a penguin a function of the sex of the animal?

Add a code chunk by hitting the button or by typing "Ctr+Alt+I":



```{r}
lm(body_mass_g ~ sex, data = d)
```


## Data preparation

```{r}
library(dplyr)
# library(palmerpenguins)

penguins_filtered <- 
  d %>%
  filter(sex %in% c("male", "female"))
```


Control+Enter -> Run one single line OR the highlighted area



```{r}
library(tidyverse)
penguins_filtered |> 
  count(sex)
```

# Cleaned data



```{r}
lm_weight_sex <- lm(body_mass_g ~ sex, data = penguins_filtered)
```


Visualize the data and the model:

```{r}
library(tidyverse)
library(ggpubr)
```


```{r}
ggdotplot(penguins_filtered,
          x = "sex",
          y = "body_mass_g",
          add = "mean")
```

Quality of the model, "goodness" of fit:

```{r}
library(easystats)
```

`r2` gives you the R-Squared:

```{r}
r2(lm_weight_sex)
```


# `y ~ x`

DV: y (metric)
IV: x (metric)

Research question: 

> body mass as a function flipper length


```{r}
lm_weight_flipper <-
lm(body_mass_g ~ flipper_length_mm, data = penguins_filtered)
```




